A partisan affair: Mapping patronage in municipal bureaucracies of Brazil

### 2021-03-25 paper/draft.tex
the recruitment of local elites

\begin{figure}[H]
    \centering
    \includegraphics[width = 6cm, height = 6cm]{figures/partisanship/plot_partisanship_by_cbo_group.pdf}
    \includegraphics[width = 6cm, height = 6cm]{figures/partisanship/plot_partisanship_by_contract_type.pdf}
    \caption{\textbf{Proportion of party-affiliated members}. On the left, breakdown of party membership by types of occupation. On the right, by type of contract.}
    \label{fig:partisan_breakdown}
\end{figure}

@@

>>>>

\begin{figure}[H]
    \centering
    \includegraphics[width = 7cm, height = 7cm]{figures/partisanship/plot_partisanship_by_cbo_group.pdf}
    \includegraphics[width = 7cm, height = 7cm]{figures/partisanship/plot_partisanship_by_contract_type.pdf}
    \caption{\textbf{Proportion of party-affiliated members}. On the left, breakdown of party membership by types of occupation. On the right, by type of contract.}
    \label{fig:partisan_breakdown}
\end{figure}

@@

Finally, party members once in the bureaucracy both earn more and stay for longer than their non-partisan counterparts. Figure \ref{fig:partisan_trajectory} illustrates these trends. Between 2003 and 2017, party members earned an average premium of around 22 percent over their non-partisan counterparts. Additionally, they have worked for longer in the bureaucracy as well, with an average 16 percent higher tenure in the municipal government. Note that while we aggregate these different figures across the years, there is widespread variation across municipalities in these different premiums. Unpacking these average values is an important avenue for future research.

\begin{figure}
    \centering
    \includegraphics[width = 7cm, height = 7cm]{figures/turnover/plot_wage.pdf}
    \includegraphics[width = 7cm, height = 7cm]{figures/turnover/plot_work_experience.pdf}
    \caption{\textbf{Partisan trajectories in the bureaucracy.} Party members earn more while working in the bureaucracy (left) and also tend to work for longer (right).}
    \label{fig:partisan_trajectory}
\end{figure}

In sum, the evidence suggests that party members concentrate in the summits of power. This reinforces priors regarding the control of bureaucratic management by politically appointed bureaucrats in a quantifiable way and in the context of a large, developing country with decentralized structure of local governance \citep{lewis2010politics}. Connecting this empirical evidence with the record that the majority of party members are coming from the upper strata of the formal economy, it becomes evident that party members are in fact occupying the upper echelons of the private and public economies, accruing important financial gains in the process and over the long run.

This raises the question of what are the broader implications of local party organizations on the compensation and incentives for its members to join. Is this simply a club good that provides access to public sector employment? And why are only members of the wealthier classes being recruited? In the next section, I explore the incentives for parties to recruit the local elites for running campaign elections. Brazil contains detailed information on party contributions that I leverage to explore the motivation for recruitment of local elites.
